### 2022-04-26 pj-7.1/response.txt
pj: first-draft response + updated paper

TODO
- write a quick thank-you and summary of changes
- add section + page numbers to the responses

> Review #45A
> ===========================================================================
>
> Paper summary (for authors)
> ---------------------------
> The paper makes a contribution to gradual typing in theory and practice. The
> subject language is Static Python, which adds a compile-time type discipline to
> Python.
>
> On the theory side, the paper describes a formal model created with an
> explorative tool for semantics modeling and prototyping. This model includes a
> static semantics and a dynamic semantics. The dynamic semantics is tested
> against a regression test suite.
>
> On the practical side, the paper describes the design of Static Python as well
> as experiments with an implementation of Static Python. This implementation
> allegedly improves the throughput of a closed source implementation of a web
> server by 3.6%. It improves the run time of several microbenchmarks
> significantly.
>
> Assessment and comments (for authors)
> -------------------------------------
> ## strengths
>
> * The main strength of the paper is clearly the favorable evaluation of a
>   particular datapoint in the design space of gradual typing.
> * The formal model enabled the discovery of problems in a industrial strength
>   implementation of Static Python, at least in its type checker.
>
> ## weaknesses
>
> * the most significant evaluation is on a closed source program
> * the microbenchmarks potentially gain more efficiency from progressive
>   primitive types as most of them are computationally intensive.
> * the theoretical evaluation is only probabilistic with unclear probability
>
> ## comments
>
> I like the paper, it is generally well readable. There are a few places, where
> I'd expect more guidance from the authors, rather than expecting the reader to
> interpolate.
>
> ### Table 1
>
> The caption is misleading. The data is the cost of creating and using
> shallow/concrete dictionary types in typed and untyped contexts.  I was
> initially confused by the 0 entries. I'd prefer you put --- to indicate that no
> casts is needed.  You should indicate that $N$ is the length of the dictionary.

Thanks, we've updated the caption, explained N, and used --- for no cast.

> ### enforcement of CheckedDict[K,V]
>
> *... gets enforced in constant-time* What if V is also a CheckedDict?

We've added an explanation of nested types (they do run in constant time).

> ### p8, 2
>
> *Nom [...] flexible point in the design space*
> Explain this point in two sentences.

Done.

> ### p8, se 2.5
>
> *...static semantics is progressive...*
> It seems like the second paragraph *Primitive types...* explains the
> progressive aspect. You should affirm that in the text.

We've expanded the first paragraph and revised the second paragraph. The first
paragraph focuses on the "normal" progressive aspects. The second focuses on
gradual typing.

> ### p9, -3
>
> *one token*
> Is this a variable size token or fixed size (if V is CheckedDict? see above)

It is always a fixed-size pointer.

> ### p10, 2
>
> *only if*
> Is that an `iff`?

No, the "only if" is intentional. An "iff" would be correct, but the paragraph
in which this sentence appears is about one direction in particular.

> ### p10, -17
>
> *Redex's random testing ... to check type soundness*
>
> That seems a bit shaky. What's the baseline / ground truth / oracle for these tests?
> In the end, all that testing only gives you a probabilistic guarantee.
> Can you give an estimate of this probability? (I mean, quantitative)

These are property-based tests. Our model defines a soundness property for terms
and asks Redex to generate random terms in the hope of finding a counterexample.
So, the soundness property is the ground truth; these random tests are all
checking soundness.

One could imagine a probabilistic guarantee about what kinds of terms the tests
explore, but the guarantee would depend on two complicated parameters: our
model's grammar and Redex's algorithm for exploring grammars (Klein & Findler,
SFP 2009). These parameters make it difficult to formulate a guarantee. On top
of that, **explaining** the guarantee to readers would be a huge challenge.

Property-based testing is shaky compared to formal verification, but it's still
a helpful tool for finding bugs.  Indeed, we found several bugs in our model
thanks to these property-based tests (none of which had implications for Static
Python itself).

> ### p10, -2
>
> *All but one...*
>
> Which soundness bug was **not** fixed and why?

All soundness bugs have been fixed at this time.

> ### p11, -3
>
> `Dyn` should be `Any` in Python.

The revised paper explains why we use the name `Dyn`.

> ### p12, fig 3, 4
>
> * `str` should be italic
> * why is `object` part of the expression language?

Fixed `str` and removed `object`.

(`object` was there as a simple instance of the `Object` type. Since it's not
needed to explain type boundary issues, we can remove it.)

> Line 6: `C.f(expr)`
> * looks like a static method call to me, is that right? That's usually trivial
>   compared with the dynamic dispatch of a real method call.
> * how do you invoke a method on an object?

Oh dear. This was not meant to be a static method call, but we were far too
aggressive when simplifying the model.

The revision uses `expr.f(expr)` instead and looks up the class type of the
receiver to type check the call.

> * how do you create an object? Is there a syntax like `C(expr)`?

Syntax for making an object was missing from the paper (not from SP or Redex!).
We've added it, thanks!

> ### fig 4
>
> Some evaluation types don't make sense
>
> * can Optional types be nested? E.g. Optional[Optional[int]]   doesn't make
>   sense to me
> * what about Optional[None] ??
> * or Optional[Dyn]??

Yes, those types can all be written. Just like surface unions (section 4.1), we
assume that Optional evaluation types get normalized in a standard way. The
normal forms for these examples are: `Optional[int]`, `None`, and `Dyn`.

> BTW, I find it confusing that surface types are `T`, but evaluation types are
> `S`, given that surface starts with s.

Good point, we've switched `S` and `T`.

> ### table 2
>
> Misleading caption: this is about the cast from Any to an evaluation type.

We have reworded the caption.

> ### p14, fig 5
>
> * rule `M-App` seems too limiting. Literally read, it's not possible to invoke
>   inherited methods because the method `f0` must be defined in class `C0` in
>   which it is defined. This rule affirms the class-method conjecture from
>   above.

We had envisioned the lookup `Env(C0)` as searching both `C0` and its parents
for all visible fields and methods.

The revised figure writes `Env*(C0)` instead and includes a note to this effect.
(The revision also uses `\Gamma` instead of `Env`)

> * the `object` rule for subtyping contradicts Table 2, which says that *objects
>   accepts any value expect primitives*. So `S0` must not be a primitive; but
>   primitives are not part of the model. This is not consistent!

We've moved the Table 2 comment about primitives to a footnote (because, as you
rightly point out, they are not part of the model).

> * given transitivity, the last rule could be simplified by omitting the
>   assumption `C2 <: C1` and changing the conclusion to `C0 <: C1`.

Thanks!

> ### p15, -16
>
> *better left to the Redex mechanization*
>
> It would be useful to show the excerpt for method calls here, as they seem to
> be nonstandard.

Method calls in Redex are standard; they work like calls in (Static) Python.
The confusion here is due to our over-simplified model in the submission, which
gave the impression that only a class can receive a method call.

> ### p24
>
> references [23] and [24] and [41] have broken DOIs

Fixed.

> * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *
>
>
> Review #45B
> ===========================================================================
>
> Paper summary (for authors)
> ---------------------------
> The contribution of the paper is the following:
>
> - The paper describes Static Python, a project developed at Instagram to add
>   types to Python code in a gradual manner, with the main objective of
>   improving performance. The project has already shown its effectiveness. The
>   authors (informally) describe both the language (the gradual type system) and
>   how this is implemented to achieve better performance.
>
> - The authors have developed an implementation in PLT redex of a subset of
>   Static Python, focusing on features related to gradual typing. Using this
>   implementation, they did an extended testing to show:
>   1) that this implementation is sound
>   2) that this implementation gives the same results of Static Python.
>
> Assessment and comments (for authors)
> -------------------------------------
> Assessment: The contributions of the paper summarized above are certainly
> useful for the PL community, so I support acceptance. The empirical evaluation
> seems to me (but I am definitely not an expert in such kind of contributions)
> well-driven and adequate. I have some reservations on the formalization,
> explained in the comments below, which should be solved/improved in the final
> version.
>
> Detailed comments:
>
> Page 3 Mechanization: you should explain better what do you mean by
> "soundness". I thought of standard soundness in the sense that "well-typed
> programs do not go stuck", but here you say "property-based soundness tests",
> what do you mean more precisely?

Section 4 in the revised paper gives more details.

In short, we started with the property "well-typed programs do not get stuck"
and tested it on thousands of terms generated by Redex.

> Section 4.1 line 4: In the formal syntax in Figure 3, Object seems not to be an
> instance of C, so class C(C) seems not to include the case C = Object.
> In the "Other expressions" you mention object fields reads and writes, but in
> the formal syntax in Figure 3 you have C.x and C.x = expr, I was expecting
> expr.x and expr.x = expr' instead.

We've made several changes in response:

- `Object` is no longer part of the grammar
- The grammar comes with a note that the set `C` must contain `Object`
- Field reads and writes use `expr` instead of `C` (example: C.x is now expr.x)

> Figure 3: it would be better to use a more abstract style with metavariables as
> it is customary now, e.g.
> prog = stmt_1, ..., stmt_n
> expr = ... [x_1:expr_1, ..., x_n:expr_n] | expr[expr'] |

We appreciate the suggestion, but prefer to keep the current grammar because it has
no substripts and uses a uniform style for the "sequence" terminals `prog` and `Env`.

> Page 13 line -7: A dynamically-typed variable -> A dynamically-typed function

No, "variable" is there on purpose because we don't know statically what a
dynamically typed variable points to. It might not be a function.

> Figure 5 rule D-App: in Figure 3 a function f seems not to be allowed to have
> type Dyn in the enviroment

This is allowed, but only indirectly. The definition of a function `f_0` can't
directly say that `f_0` has type Dyn, but it can declare the type `Dyn -> Dyn`
and then apply `f_0(f_0)` to assign the dynamic type to a function.

(Note that `x` and `f` are undifferentiated variable names. The model uses `x`
 in some definitions and `f` in others as hints to readers, but it's fine to
 name a function `x_2` and so on.)

> Also, Object seems to be the top type, hence also Bool and Int are seen as
> classes, right?

Yes, Bool and Int are technically classes. The paper says little about this,
but the Redex model (and Static Python) give them full support as classes.

> In rules D-Set and CD-Set, the type None seems to be used as a "Unit" type. The
> fact that these two types are identified seems strange.

Yes it is a bit strange, but it helps keep the model small for the paper.
(Really, the best choice would be to move these rules into a typing judgment
for **statements** rather than expressions. That's what the Redex model does.)

> In subtyping rules, which is the relation between two Optional types?

Fixed, thank you for catching this omission.

> Why do you use metavariable Env_0 rather than just Env in the rules?

We've removed the subscript.

> page 15 line 4: I would not say "upcast" since there is no explicit cast here.

Ok. We have removed this explanatory sentence.

> My main criticism about your formalization of the typing rules is that, in my
> opinion, they should include generation of code annotated with casts.
> Moreover,it should be made clear that they are algorithmic. I mean, what I
> expect is that there is an algorithm which, taken an expression expr and an
> environment Env, produces a type (or a typing error) and the corresponding
> annotated code, inserting casts. I am sure that this is what happens in your
> PLT Redex implementation, and I think that this should be reported.

The typing judgment in Redex is algorithmic, but the simple judgment in the
paper is not. Section 4.2.2 in the revised paper makes this clear.

> Section 4.3: I would perhaps eliminate this section. As it is, it says almost
> nothing.

We disagree. The section is a proof sketch that ill-typed values cannot "sneak
in" through the dynamic type.

> Section 6.3 I miss a conclusion/summary of the analysis here.

We've added a conclusion section.

> * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * *
>
>
> Review #45C
> ===========================================================================
>
> Paper summary (for authors)
> ---------------------------
> This paper presents the design of Static Python, which introduces
> gradual typing for Python.  It is already in use by engineers at
> Instagram.  One of the notable features in the design is that a
> programmer can have some control over the choice of what kind of
> invariant can be enfored by types: untyped (no guarantee), shallow (in
> the sense that types guarantee only the top-level type constructor),
> and concrete (where full type information is relevant).  The use of
> concrete types promotes code optimization, which can be also an
> incentive to programmers.
>
> The paper presents an overview of the language and its run-time
> system, followed by the description of the formal type system.  The
> type system is formalized with PLT Redex.  Although no formal
> metatheory is developed, the model has been checked by Redex's random
> testing tools and test suites.  Subtleties in the language features
> that are not modeled but implemented are discussed, too.  The section
> on evaluation reports that 3.7% imprevement in CPU efficiency.
>
> Assessment and comments (for authors)
> -------------------------------------
> The language design mixing concrete and shallow types is interesting.
> It seems that the language has been accepted by engineers and used for
> a large software project.  Nevertheless I have a few reservations to
> accept the paper in its current form.
>
> The paper puts emphasis on "gradual soundness" at the beginning but,
> after reading the paper, I failed to see what it really was.  The
> phrase is never mentioned on page 3 and afterwards.  This is
> unfortunate because this is the phrase used even in the title!

We've updated the paper to mention "gradual soundness" in the following
sections: 2.3, 2.5, 6.3, and 8.

> It's not clear how faithfully the formal model presented in Section 4
> models the implementation.  In particular, C-Sub allows the type of an
> expression can be changed to Dyn (using the consistent subtyping) and
> Matr allows conversion from Dyn to any non-Dyn type.  So, the type
> system seems to give any expression (that references only declared
> variables, functions, classes) any type.  However, apparently, some
> expressions are rejected by the implementation.  I have no idea what's
> really going on.  (I tried to access the Redex model but it wasn't
> accessible.)

We've added a disclaimer near the top of section 4.2.2 to say that the
paper presents declarative typing rules whereas the Redex model implements a
typing algorithm.

> The distinction between surface types and evaluation types is
> interesting but its implication is not so clear.  Are type arguments
> to Dict completely ignored by the type system?  If so, what is the
> point of writing type arguments in a program?  I may misunderstand
> something and am wondering if my question is related to the paragraph at
> the bottom of p.11, "we omit the surface typing judgment, which is a
> kind of linter that merely scans typed code for logical errors".  What
> are the logical errors mentioned here?  If type arguments to Dict are
> completely ignored, an alternative design might have been like this:
> instead of providing two classes Dict and CheckedDict, a programmer
> could use the raw type Dict and fully parameterized types Dict[S0,S1].

The discussion is intentionally vague because Static Python does not have firm
guidelines for what the surface type checker (linter) should do. For Dict,
it currently ignores the type arguments so the only reasons to write them are
for documentation and for other type checkers (Pyre, mypy) to check them.
In the future, though, Static Python may do more surface checks.

Using the raw type Dict and parameterized type Dict[S0, S1] seems problematic
because these types describe different data structures. The types look similar,
but a value with type Dict[S0, S1] cannot be used where a Dict is expected.

> The related work section is light.  It cites relevant papers and
> describes what they discuss but doesn't really compare with the
> present work in depth.

We have expanded the related work.

> In summary, I'd recommend that the paper be revised by:
> * clarifying gradual soundness,
> * clarifying the relationship between the model and the implementation,
> * clarifying the role of surface types,
> * enhancing the discussion on related work.
> It would be appreciated if you make the code of the Redex model accessible.

We apologize that the Redex model was inaccessible. For convenience, here is
the link:

  <https://github.com/brownplt/insta-model>

> Detailed comments:
>
> Table 1: Why does a constructor call for CheckedDict pay the cost in
> O(N)? (What does N stand for?)

N is the number of keys and values in the table (we have updated the table
caption to explain). The constructor needs to validate each element.

> * It seems that efficient run-time checking depends on the fact that
>   the set of type instantiations CheckedDict[S0, S1] is finite and
>   known at compile-time.  How is it achieved?  (Is it because there
>   are no user-defined generic classes?)  My guess above may be wrong.
>   In that case, it's not very clear how tags are implemented.

Section 2.3 and 3 in the revision say more about the "how." Static Python
implements these tags as pointers into a run-time table of all generic
instantiations that have appeared.

> p.12: "Evaluation types S are a subset of the surface types".
> Is "Dict", which is an evaluation type, a surface type?

No, Dict is not a surface type in the model. We have rephrased "subset" to
avoid this confusion.